AttCRISPR : an spacetime interpretable model for sgRNA efficiency prediction

### 2020-10-18 Paper/main.tex
10-17 data
10-18 paper

As Figure~3\vphantom{\ref{fig:03}} show, Given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{matrix}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{matrix}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.
\subsubsection{Method and attention in temporal domain}
As Figure~4\vphantom{\ref{fig:04}} show, temporal domain part of AttCRISPR relies on the recurrent neural network (or its variant). 
As previously mentioned, we mapping each one-hot vector into a dense real-valued high-dimensional space follow the Equation~(\ref{eq:08}), which generates the embedded matrix $X_e$.
And we regard $X_{e}$ as a sequential data, or temporal data.
Recurrent neural network (or its variant) has showed outstanding performance in the tasks with temporal data (for instance, natural language processing, sequential recommendation). 
That's why we prefer to use it to extract potential temporal features. 
To be precise, we prefer the architecture of encoder-decoder which has been proven to be effective in Seq2Seq task. 
Two main differences we have to face are that sgRNA is not a natural language in the traditional sense, and we don't have to translate it to other sequence. 
To accommodate them, the embedded matrix $X_e$ is used as input of both the encoder and decoder, and the sequence of decoder is to build the first-order preference of sgRNA sequence $\tilde{A}$. 
As mentioned above, the predicted value $y$ should satisfy Equation~(\ref{eq:05}).

@@

\left\{\begin{matrix}

@@

\end{matrix}\right.\label{eq:14}

@@

\section{Results}

We designed experiments to address the following questions:
\begin{itemize}
    \item 
    Whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions make by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

>>>>

As Figure~3\vphantom{\ref{fig:03}} show, Given an one-hot map $X_{oh}$, spatial attention module generating spatial attention matrix 
$A_s\in\mathbb{R}^{l\times 4}$ with the same as $X_{oh}$ in shape. 
Each element of $A_s$ is constrained to a range of zero to one, implemented by a sigmoid function, which reflects the importance of the corresponding elements of $X_{oh}$.
The overall spatial attention process can be summarized as:
\begin{equation}
\left\{\begin{array}{l}
X_{mc} = f^{3\times4}(X_{oh})
\\A_s = \sigma(f^{3\times2}([AvgPool(X_{mc});MaxPool(X_{mc})]))
\\X_{rf} = A_s\otimes X_{oh}
\end{array}\right.\label{eq:11}
\end{equation}
where $f^{p\times q}$ represents a convolution operation with the filter size of $p\times q$, 
$p,q\in\mathbb{Z}^{+}$, $X_{mc}$ is a multi-channel map generated by $X_{oh}$, $\sigma$ denotes the sigmoid function, 
AvgPool denotes the average-pooling operation, MaxPool denotes the max-pooling operation, $\otimes$ denotes element-wise multiplication.
Spatial attention matrix $A_s$ formally conforms to our proposed definition of first-order preference (each element of $X_{oh}$ is multiplied by the corresponding element of $A_s$), 
in other word, element of $A_s$ reveal how important the corresponding elements in $X_{oh}$ is. 
We think it can reveal the preference of the scoring function at each position. 
For instance, follow the encoding rules above, train spatial domain part of AttCRISPR with the WT-SpCas9 dataset. 
Then take the average of all spatial attention matrix, and the element in the first row and third column are closer to 1, which means when calculate the final score, 
G typically may have a important contribution at first position within sgRNA sequence. 
In fact, this corresponds to some early studies concerning the Human (hU6) promoter, which is believed to require G as the first nucleotide of its transcript \citep{cong2013multiplex,jinek2012a,mali2013rnaguided}.
\subsubsection{Method and attention in temporal domain}
As Figure~4\vphantom{\ref{fig:04}} show, temporal domain part of AttCRISPR relies on the recurrent neural network (or its variant). 
As previously mentioned, we mapping each one-hot vector into a dense real-valued high-dimensional space follow the Equation~(\ref{eq:08}), which generates the embedded matrix $X_e$.
And we regard $X_{e}$ as a sequential data, or temporal data.
Recurrent neural network (or its variant) has showed outstanding performance in the tasks with temporal data (for instance, natural language processing, sequential recommendation). 
That's why we prefer to use it to extract potential temporal features. 
To be precise, we prefer the architecture of encoder-decoder which has been proven to be effective in Seq2Seq task. 
Two main differences we have to face are that sgRNA is not a natural language in the traditional sense, and we don't have to translate it to other sequence. 
To accommodate them, the embedded matrix $X_e$ is used as input of both the encoder and decoder, and the sequence of decoder is to build the first-order preference of sgRNA sequence $\tilde{A}$. 
As mentioned above, the predicted value $y$ should satisfy Equation~(\ref{eq:05}).

@@

\left\{\begin{array}{l}

@@

\end{array}\right.\label{eq:14}

@@

\section{Results}
\begin{figure*}[!tpb]%figure4
    \centerline{\includegraphics[width=178mm]{baselinewithoutbiofeat.eps}}
    \caption{wait2write}\label{fig:06}
    \end{figure*}
We designed experiments to address the following questions:
\begin{itemize}
    \item 
    In the absence of hand-crafted biological features, whether the stacking of method in spatial domain and temporal domain can get better performance then use these method alone? 
    $\rightarrow$ Section~\ref{section:stacking}
    \item 
    How does AttCRISPR perform compared to current state-of-the-art methods, covering both conventional machine learning and deep-learning models? 
    $\rightarrow$ Section~\ref{section:comparison}
    \item 
    How can researcher understand the decisions make by AttCRISPR locally and globally, based on attention mechanisms? 
    $\rightarrow$ Section~\ref{section:interpretability}
    \vspace*{1pt}
\end{itemize}

@@

Text
\begin{table}[!tpb]
    \processtable{wait2write\label{Tab:02}} 
    {\begin{tabular}{@{}llll@{}}\toprule
        Method & WT-SpCas9 & eSpCas9(1.1) & SpCas9-HF1\\\midrule
        XGBoost & 0.845 & 0.831 & 0.818\\
        MLP & 0.842 & 0.846 & 0.844\\
        CNN & 0.846 & 0.831 & 0.834\\
        RNN & 0.856 & 0.849 & 0.851\\
        TAC & 0.857 & 0.844 & 0.851\\
        SpAC & 0.862 & 0.854 & 0.857\\
        \textbf{EnAC} & \textbf{0.868} & \textbf{0.859} & \textbf{0.862}\\
        CRISPRpred(SEQ) & 0.838 & 0.830 & 0.821\\
        \botrule
    \end{tabular}}\footnotesize\setlength{\parindent}{2em}{\emph{Note}: wait2write}
\end{table}
